An Efficient Calculus for Generalised Parikh Images of Regular Languages

### 2022-06-21 main.tex
Add duels
Mainly work on introduction

\contents{
  \item introducera homomorfismen
  \item start with length
  \item räkna bara a och b, strunta i c
  \item ändra till >= för att illustrera splitting
  }

Consider the two automata $\SomethingCSomething{}$
(Figure~\ref{fig:something-c-something}) and $\AcaOrBc{}$
(Figure~\ref{fig:aca-or-bc}). Assume we want to find a value in the Parikh image
of their product where $\TransitionCount{a} > \TransitionCount{b}$, that is find
a count of characters in a string accepted by both automata where there are more
a:s than b:s.

The na\"ive approach would be to first compute the product of the automata and
then compute the Parikh image of its product. However, we can take advantage of
the constraint to prune the automata before computing the product and help
reduce the blow-up. To give an intuition for the core idea of \Calculus{}, we
will annotate the transitions of $\SomethingCSomething$ and $\AcaOrBc$ with
variables illustrating how many times they are taken, and perform informal
reasoning on them corresponding to the reasoning performed by our calculus.

\begin{figure}[ht]
  \begin{minipage}[b]{0.75\linewidth}
  \centering 
    \includegraphics[width=\textwidth]{aca_or_bc}
    \caption{An automaton recognising the regular expression
    $\mathtt{ac^*a|b(bb)^*c}$.}\label{fig:aca-or-bc}

@@

Note that we have already performed some reasoning in the starting figures -- we
use the same symbol for the number of times each of the upper transitions of
$\AcaOrBc$ are taken, since we know that they must be taken the same number of
times, and we introduced the constant $1$ for the middle transition in
$\SomethingCSomething$, since it must always be taken precisely once to reach
the accepting state. We set the initial flow into the starting state to 1 and
use linear equations to enforce balanced flow throughout the automaton to
determine the relations between the counting variables on each transition (e.g.
$1 + {l_b}' = l_a + l_b$, or $l_a + l_c' = 1$). Using standard linear equation
reasoning loses additional variables and gives us the annotations in
Figure~\ref{fig:example-simplify-1}.

\begin{figure}[h]
  \begin{minipage}[b]{0.75\linewidth}
  \centering 
    \includegraphics[width=\textwidth]{aca_or_bc_simplified}
    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
    simplification of their equations.}\label{fig:example-simplify-1}
  \end{minipage}
  \end{figure}

  We now need to relate the number of times a transition was taken to the
  character counts we are really interested in. Note that this also puts the two
  automata into contact with each other, before any product has been computed.
  Some consideration leads us to the definition that the number of occurrence of
  a each character corresponds to the sum of usages of transitions containing
  it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c$, and
  $\TransitionCount{b} = 2 l_b' + l_c'$.\footnote{For the attentive reader: in this
  example, we leave the catch-all transitions of $\SomethingCSomething$
  unconstrained since they will be taken care of by the product and would
  complicate the reasoning since we cannot yet know which character they will
  correspond to in the final product. In other words, this is a safe
  under-approximation that only limits our ability to reason before
  computing the product.}

@@

\begin{figure}[h]
  \begin{minipage}[b]{0.75\linewidth}
  \centering 
    \includegraphics[width=\textwidth]{aca_or_bc_simplified_2}
    \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
  \end{minipage}
  \end{figure}

  Plugging in the same inequality again on the same principle, we get $2(1 -
  l_c') > l_c' \equiv 1 > l_c'$ since we are working with integers, and
  therefore that $l_c' = 0$. This gives us the even smaller automaton of
  Figure~\ref{fig:example-simplify-3}.

@@

\subsubsection{An Example}
We apply Equation~\ref{eq:generate-parikh} to the automaton of Figure~\ref{fig:len-branch}. We then get the quantified Presburger formula seen in Equation~\ref{eq:parikh-formula}. As before, we omit the implicit existential quantifiers for brevity, along with the constranints requiring all variables to be nonnegative. We also omit the $\land$-clauses between each line. In other words, in this example, only the variables $\LetterVar_a, \LetterVar_b, \LetterVar_c, \LetterVar_d, \LetterVar_e$, and $\LetterVar_f$ are free, and corresponds to the counts of the respective subscript letter. Note that since they here uniquely identify a transition due to the one-to-one mapping between character labels and transitions in the example, the first clause can be trivially eliminated by substituting $\LetterVar_a$ for $\TransitionVar_{\FromLabelTo{0}{a}{1}}$, etc. This is not generally the case.

\begin{equation}
  \begin{aligned}
  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} \land \LetterVar_{d} = \TransitionVar_{\FromLabelTo{2}{d}{3}} \land \LetterVar_{e} = \TransitionVar_{\FromLabelTo{2}{e}{2}} \land \LetterVar_{f} = \TransitionVar_{\FromLabelTo{1}{f}{3}}\\
  &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
  & \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{0}{a}{1}} = 1 \land  \TransitionVar_{\FromLabelTo{1}{f}{3}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
  & \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{2}{d}{3}} - \TransitionVar_{\FromLabelTo{1}{f}{3}} = 0\\
  &\TransitionVar_{\FromLabelTo{0}{a}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} > 0 \implies \StateVar_{2} > 0\\
  &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} > 0 \implies \StateVar_{2} > 0\\
  &\TransitionVar_{\FromLabelTo{1}{f}{3}} > 0 \implies \StateVar_{3} > 0 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} > 0 \implies \StateVar_{3} > 0\\
  &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{0}{a}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} \geq 1 \land \StateVar_{2} \geq 1\right)\\
  &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{f}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{2}{e}{2}} \geq 1 \land \StateVar_{2} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{d}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
  &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
  \end{aligned}
  \label{eq:parikh-formula}
  \end{equation}

@@

@@

  $$
  \BindingSum(\Automaton_1 \times \Automaton_2, \Filter) = \bigcup
  \begin{aligned}
  & \Set{
    \Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_1}
  }, \\ 
  & \Set{
    \Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_2}
  }
  \end{aligned}
$$

@@

\begin{figure}
  \includegraphics[width=0.75\linewidth]{graphs/\commit-duels-scatter.pdf}
  \caption{The pairwise runtime of each instance for instances solved by any backend. Timeouts are reported as a runtime of \RuntimeTimeout.}
  \label{fig:duels}
\end{figure}

>>>>

To illustrate the key features of our approach, we start with two regular
languages, a commutative homomorphism $f$ on them that we care about, and a side
constraint for which we want to find values in $f$'s image over the intersection
of the wto languages. We will use the same two regular languages, recognised by
two automata; $\SomethingCSomething{}$ (Figure~\ref{fig:something-c-something})
and $\AcaOrBc{}$ (Figure~\ref{fig:aca-or-bc}) for both examples but change the
constraint and homomorphism $f$.

\subsubsection{Counting string lengths}

To keep things simple, in this first example we
let $f$ be the function that gives the length of a given string, e.g.
$f\left(\text{curious and curiouser still}\right) = 27$, and $f\left(\text{homo}\right) +
f\left(\text{morphism}\right) = f\left(\text{homomorphism}\right) = 12$.

\begin{figure}[h]
  \begin{minipage}[b]{0.75\linewidth}
  \centering 
    \includegraphics[width=\textwidth]{aca_or_bc}
    \caption{An automaton recognising the regular expression
    $\mathtt{ac^*a|b(bb)^*c}$.}\label{fig:aca-or-bc}

@@

The na\"ive approach would be to first compute the product of the automata and
then compute the length of its product. However, we can take advantage of the
both constraints and definitions of $f$ to prune the automata before computing
the product and help reduce the blow-up. To give an intuition for the core idea
of \Calculus{}, we will annotate the transitions of $\SomethingCSomething$ and
$\AcaOrBc$ with variables illustrating how many times they are taken, and
perform informal reasoning on them corresponding to the reasoning performed by
our calculus.

\begin{figure}[h]
  \centering 
    \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified}
    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
    simplification of their equations.}\label{fig:example-simplify-1}
  \end{figure}

Note that we have already performed some reasoning in the starting figures -- we
use the same symbol for the number of times each of the upper transitions of
$\AcaOrBc$ are taken, since we know that they must be taken the same number of
times, and we introduced the constant $1$ for the middle transition in
$\SomethingCSomething$, since it must always be taken precisely once to reach
the accepting state. We use standard graph-theoretical flow reasoning to set the
initial flow into the starting state to 1 and use linear equations to enforce
balanced flow throughout the automaton to determine the relations between the
counting variables on each transition (e.g. $1 + {l_b}' = l_a + l_b$, or $l_a +
l_c' = 1$). Strictly speaking, this is where the constant number $1$ in
Figure~\ref{fig:something-c-something} comes: from $1 + r_\Sigma = x + r_\Sigma$
solved for $x$. Using analogous linear equation reasoning loses additional
variables and gives us the annotations in Figure~\ref{fig:example-simplify-1}.

We sum the individual transition variables up to get the length we are actually
interested in, e.g. the image of $f$ on $\SomethingCSomething =
\exists_{r_\Sigma, r_\Sigma'} \: r_\Sigma + 1 + r_\Sigma'$. For
$\SomethingCSomething{}$, this is all we need to produce the image since all
transitions are always accessible from any path from the initial state to the
accepting state.

The story gets more complicated with $\AcaOrBc{}$, where we have a choice of two
paths, and where that choice affects the possibility of taking either of the two
loops. In that case (letting the existential quantifiers from now on be
implicit), we would have the image $2l_a + l_c + l_c' + l_b + l_b'$. If we apply
standard linear reasoning on the equations (e.g. $1 + {l_b}' = l_a + l_b$, or
$l_a + l_c' = 1$), we can reduce the number of bound ($\exists$-quantified)
variables somewhat, obtaining the relations in
Figure~\ref{fig:example-simplify-1}, and the corresponding over-approximation on
the $f$-image $2 + l_c + 2l_b'$. It is an over-approximation because it only
considers the number of times each transition is taken and the existence of a
path from an initial to an accepting state, but omits the connectivity of that
path in the presence of loops. For example, no valid path would have both $l_c >
0$ and $l_b' > 0$, but no such restriction exists in the formula above.

\begin{figure}[h]
  \centering 
    \includegraphics[width=0.75\linewidth]{something_c_something_length_simplified}
    \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
  \end{figure}

Still, this approximate solution allows us to perform further reasoning. Since
we know that the length of a string accepted by the intersection of the two
automata's languages must be the same, we also know that their images (string
lengths) must be the same under $f$, e.g. that $r_\Sigma + 1 + r_\Sigma' = 2 +
l_c + 2l_b' \iff r_\Sigma + r_\Sigma' = 1 + l_c + 2l_b'$. This in turn allows us
to replace one additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' -
r_\Sigma'$ (see Figure~\ref{fig:example-length-reduced}). Note that the two
automata are now in direct contact. Any reasoning performed on one of them could
be propagated to the other.

Since we cannot take a transition a negative number of times, we have an
implicit constraint on any transition variable that it must be at least $0$.
This means that we have $1 + l_c + 2l_b' - r_\Sigma'$, implying an upper bound
on $r_\Sigma'$: $r_\Sigma' \leq 1 + l_c + 2l_b'$.

We talked initially about having a constraint on the image (or conversely, on
the strings of the language), so let us add one. We are only interested in
strings of odd length, e.g. $2 \nmid  f(\AcaOrBc \times \SomethingCSomething)$.
This means that: $2 \nmid 2 + l_c + 2l_b' \iff 2 \nmid l_c$. To encode this condition, we simply replace $l_c$ with $2k + 1$ with an existentially quantified integer $k \geq 0$.

This is as far as we can get with this reasoning. We must now choose: either we
compute the product and continue our reasoning, or we perform a case split. Let
us try a case split in order to put off computing a potentially large product.
We pick an early transition in the largest automaton with loops, in our case
$\FromLabelTo{start}{a}{sawA}$ of $\AcaOrBc$. This creates two cases: $1-l_c' =
0$ and $1-l_c' > 0$. We start with the first case, and after propagating the now known values and removing any transition that becomes zero, we get the automaton of Figure~\ref{fig:aca-or-bc-length-split-1}.

\begin{figure}[h]
  \centering 
    \includegraphics[width=0.75\linewidth]{aca_or_bc_length_split_1}
    \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{start}{a}{sawA}$ is not used.}\label{fig:aca-or-bc-length-split-1}
  \end{figure}

  We immediately notice that the self-loop of state $sawA$ is now disconnected
  and must be removed. This condition can be detected efficiently using standard
  forwards/backwards reachability from the initial and accepting states
  respectively. In other words, $2k+1 = 0$ for this case and is implied by the splitting constraint. However, no choice of integer $k$  will satisfy this equation. Therefore, we close this branch and proceed with the other one -- where $1-l_c' > 0$. Note that we do so without ever computing a potentially expensive product of automata.

  \begin{figure}[h]
    \centering 
      \includegraphics[width=0.75\linewidth]{aca_or_bc_length_split_2}
      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{start}{a}{sawA}$ is used.}\label{fig:aca-or-bc-length-split-2}
    \end{figure}

Plugging in the assumption $1-l_c' > 0$ (and its consequence that $l_c' = 0$),
we also get a reduced variant of $\AcaOrBc$, as seen in
Figure~\ref{fig:aca-or-bc-length-split-1}. Note that $sawB$ is no longer
backwards-reachable from $final$ and therefore dead, along with both transitions
into it. We add the conclusion $l_b' = 0$, which severs both transitions into
the state and removes it from our automaton. Following this we are ready to
compute the product, which can be seen in Figure~\ref{fig:length-split-product}.

\begin{figure}[h]
  \centering 
    \includegraphics[width=0.75\linewidth]{length_split_product}
    \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
    case splitting and removal of dead states.}\label{fig:length-split-product}
  \end{figure}

Before continuing, we need to tie the inequalities for the transitions of the
automata of the product to the new transition counting variables we introduced
in the product automaton to ensure that they are consistent. In our case, we obtain the following equations from the principle that a transition variable in either automaton must be equal to the sum of variables in the resulting edges of the product, which gives the following two interesting inequality from $\AcaOrBc{}$'s transition $\FromLabelTo{sawA}{c}{sawA}$: $2k + 1 = r_1 + r_2 + 1$

Note how this conserves the requirement that the number of loop iterations be
odd even when the same original loop appears three times in the product, twice
as a self-loop and once as a regular state transition. Using the same principle
as before for computing the length by adding up the transition counters, we
obtain the quantified formula $1 + r_1 + 1 = r_2 + 1 = r_1 + r_2 + 3 = 2k + r_1
+ 3$ for the length. Eliminating the final quantifier gives us the image of
$f(\AcaOrBc \times \SomethingCSomething)$ modulo odd lengths: $3 + 2k$, for some
integer $k$.

\subsubsection{Producing the Parikh image}

Now assume that rather than the length (e.g. the count of the characters) we
want to find the whole Parikh image of the intersection of $\AcaOrBc{}$ and
$\SomethingCSomething{}$. In other words, now $f$ is an object (usually a
vector) counting how often each character occurs in a string. Since extracting
values from a vector is tiresome, we will use the shorthand notation
$\TransitionCount{a}$ to refer to the number of letters a that appear. E.g. $\TransitionCount{a}(\text{curious and curiouser still}) = 1$. Note that we still have the property that $\TransitionCount{o}(\text{homomorphism}) = \TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1 = 3$. If you treat the addition as element-wise vector addition, this approach works for the vector version as well.

For our constraint, we will use $\TransitionCount{a} > \TransitionCount{b}$,
that is find a count of characters in a string accepted by both automata where
there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from Figure~\ref{fig:example-simplify-1}.

  We now need to relate the number of times a transition was taken to the
  character counts we are really interested in, as opposed to just the length in
  the previous example. Note that this, like the length calculus before, puts
  the two automata into contact with each other, before any product has been
  computed. Some consideration leads us to the definition that the number of
  occurrence of each character corresponds to the sum of usages of transitions
  containing it, e.g. relations of the kind $\TransitionCount{a} = 2 - 2l_c$,
  and $\TransitionCount{b} = 2 l_b' + l_c'$.\footnote{For the attentive reader:
  in this example, we leave the catch-all transitions of $\SomethingCSomething$
  unconstrained since they will be taken care of by the product and would
  complicate the reasoning since we cannot yet know which character they will
  correspond to in the final product. In other words, this is a safe
  under-approximation that only limits our ability to reason before computing
  the product.}

@@

\begin{figure}[h]
  \centering 
    \includegraphics[width=0.75\linewidth]{aca_or_bc_simplified_2}
    \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
  \end{figure}

  Plugging in the same inequality again on the same principle, we get $2(1 -
  l_c') > l_c' \iff 1 > l_c'$ since we are working with integers, and
  therefore that $l_c' = 0$. This gives us the even smaller automaton of
  Figure~\ref{fig:example-simplify-3}.

@@

\subsubsection{An Example}
We apply Equation~\ref{eq:generate-parikh} to the automaton $\AcaOrBc{}$ of Figure~\ref{fig:aca-or-bc}. We then get the quantified Presburger formula seen in Equation~\ref{eq:parikh-formula}. As before, we omit the implicit existential quantifiers for brevity, along with the constranints requiring all variables to be nonnegative. We also omit the $\land$-clauses between each line. In other words, in this example, only the variables $\LetterVar_a, \LetterVar_b, \LetterVar_c, \LetterVar_d, \LetterVar_e$, and $\LetterVar_f$ are free, and corresponds to the counts of the respective subscript letter.

\begin{equation}
  \begin{aligned}
  \ParikhMap(\Automaton):=\:
  &\LetterVar_{a} = \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{1}{a}{3}} \land \LetterVar_{b} = \TransitionVar_{\FromLabelTo{0}{b}{2}} + \TransitionVar_{\FromLabelTo{2}{b}{1}} \land \LetterVar_{c} = \TransitionVar_{\FromLabelTo{1}{c}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}}\\
  &\FinalStateVar_{3} > 0 \implies \StateVar_{3} > 0\\
  & \TransitionVar_{\FromLabelTo{0}{a}{1}} + \TransitionVar_{\FromLabelTo{0}{b}{2}} - 0 = 1 \land  \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{b}{1}} - \TransitionVar_{\FromLabelTo{0}{a}{1}} = 0\\
  & \TransitionVar_{\FromLabelTo{2}{b}{1}} + \TransitionVar_{\FromLabelTo{2}{c}{3}} - \TransitionVar_{\FromLabelTo{0}{b}{2}} = 0 \land \FinalStateVar_{3} + 0 - \TransitionVar_{\FromLabelTo{1}{a}{3}} - \TransitionVar_{\FromLabelTo{2}{c}{3}} = 0\\
  &\TransitionVar_{\FromLabelTo{0}{a}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} > 0 \implies \StateVar_{2} > 0\\
  &\TransitionVar_{\FromLabelTo{1}{c}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} > 0 \implies \StateVar_{3} > 0\\
  &\TransitionVar_{\FromLabelTo{2}{b}{1}} > 0 \implies \StateVar_{1} > 0 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} > 0 \implies \StateVar_{3} > 0\\
  &\StateVar_{0} > 0 \implies \left(\StateVar_{0} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{0}{a}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{0} = \StateVar_{2} + 1 \land \TransitionVar_{\FromLabelTo{0}{b}{2}} \geq 1 \land \StateVar_{2} \geq 1\right)\\
  &\StateVar_{1} > 0 \implies \left(\StateVar_{1} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{1}{c}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{1} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{1}{a}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
  &\StateVar_{2} > 0 \implies \left(\StateVar_{2} = \StateVar_{1} + 1 \land \TransitionVar_{\FromLabelTo{2}{b}{1}} \geq 1 \land \StateVar_{1} \geq 1\right) \lor \left(\StateVar_{2} = \StateVar_{3} + 1 \land \TransitionVar_{\FromLabelTo{2}{c}{3}} \geq 1 \land \StateVar_{3} \geq 1\right)\\
  &\StateVar_{3} > 0 \implies \FinalStateVar_{3} = 1
  \end{aligned}
  \label{eq:parikh-formula}
  \end{equation}

@@

  \contents{\item FIXME: this needs rewriting for the new automaton}

@@

  $$
  \BindingSum(\Automaton_1 \times \Automaton_2, \Filter) = \bigcup
  \begin{aligned}
  & \Set{ 
    \left<\Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State_R}}} \Filter(\Transition')\right>
  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_1} } , \\ 
  & \Set{
    \left<\Filter(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State_L, \State'}}} \Filter(\Transition')\right> \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_2}
  }
  \end{aligned}
$$

@@

\begin{figure}
  \begin{minipage}[b]{0.75\linewidth}
    \centering 
      \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-baseline-scatter}
      \caption{Runtime duel: \Calculus{} versus baseline.}
      \includegraphics[width=\textwidth]{graphs/\commit-duels-lazy-nuxmv-scatter}
      \caption{Runtime duel: \Calculus{} versus \Nuxmv.}
    \end{minipage}  
  \caption{The pairwise runtime of each instance for instances solved by any backend. Timeouts are reported as a runtime of \RuntimeTimeout.}
  \label{fig:duels}
\end{figure}
